---
title: "Class11"
author: "Celline Kim"
format: gfm
---

## Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale. So, you processed about
~230 samples and did the normalization on a genome level. Now, you want to find whether
there is any association of the 4 asthma-associated SNPs (rs8067378…) on ORMDL3
expression.

How many samples do we have? 

>Q13. Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes. 

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

We have 462 samples
```{r}
table(expr$geno)
```

```{r}
summary(expr)
```
There is a median of 25.116
```{r}
library(ggplot2)
```

>Q14. Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

Let's make a boxplot
```{r}
ggplot(expr)+ aes(geno, exp, fill=geno)+
  geom_boxplot(notch=TRUE)
```

The SNP does effect the expression of ORMDL3. We can see that having a G/G in this location is associated with having a reduced expression of this gene. 